---
output:
  pdf_document: default
  html_document: default
---
#---
#title: "Homework Task 2 -- Basics in R"
#date: 'Due: Thursday, May 2 by 08:59 CET'
#author: "Akshata Acharya" "Friedrich Heitzer" "Milad Rouygari"
#output:
#  html_document
---

# Instructions
#* Work together within your group.
#* Make sure you have R and RStudio installed. If you are an advanced user and aren't using RStudio, make sure you have `rmarkdown` in order to 'knit' the HTML output.
#* Download [this zip file](https://michael-franke.github.io/XPLab_2019/homework/XPLab_2019_HW2.zip). It contains an .Rmd file that you will edit and submit.
#* Open the .Rmd file in RStudio.
#* Fill in your names in the 'author' heading.
#* Fill in the required code and answers.
#* 'Knit' the document (`ctrl/cmd` + `shift` + `K` in RStudio) to produce a HTML file.
#* Push both the files to your group's Git repository before the deadline.
#* Using Github Pages, make a page that displays the knitted html file.

# Suggested readings
* [R4DS](http://r4ds.had.co.nz/) (R for Data Science).

# Required R packages
* `tidyverse` (or `ggplot2`, `dplyr`, `purrr`, `tibble`)

```{r}
set.seed(123) # this ensures that any psuedo-random computations are always the same
```

___

# 1. Installing and running R

a.
---

Your first task is simply to show that you have been able to install
and run R and R Markdown.  You don't have to change this code, just
uncomment it. Then the correct output will automatically appear when
you 'knit' the document.


```{r tidy = FALSE, message = FALSE, warning = FALSE, error = FALSE}
# UNCOMMENT THE CODE

#R.version

sessionInfo()
```

*Which version of R are you running? On which platform are you running it?**

*ANSWER:*

#Version 3.5.1 (2018-07-02)


b.
---

Install the package `tidyverse`. Don't install it in the code
below. Instead, install it through the console. Then write code below
to load the package and show the `sessionInfo` again.


```{r tidy = FALSE, message = TRUE, warning = TRUE, error = FALSE}

#YOUR CODE HERE
# install.packages("tidyverse")
library(tidyverse)
sessionInfo()
```

**Which version of `tidyverse` do you have installed?**

*ANSWER:*
#tidyverse 1.2.1


___

# 2. Tibbles (tidy tables)


The ToothGrowth data set comes with base R. You can read about this data set
by running `?ToothGrowth` in the console. It is a data frame. In this
course, we prefer to use tibbles (tidy tables) instead of data frames.

a.
---

Convert the `ToothGrowth` data frame into a tibble using `as_tibble()`. Put
this in a new variable called `iris_tibble`. Then print the tibble
using the `print()` function.

```{r tidy = FALSE, message = FALSE, warning = FALSE, error = FALSE}

iris_tibble = as_tibble(ToothGrowth)
print(iris_tibble)


```



**Which data type is the variable "supp"?**

*ANSWER:*
#[1] "integer

 typeof(iris_tibble$supp)


b.
---

Starting from the complete `ToothGrowth` data set, filter only the guinea pigs
that were give orange juice.

```{r tidy = FALSE, message = FALSE, warning = FALSE, error = FALSE}

# YOUR CODE HERE
OJ = subset(ToothGrowth, ToothGrowth$supp == "OJ")
OJ %>% filter(supp == "OJ")

```

**How many datapoints (i.e. guinea pigs) are left?**

*ANSWER:*

30



c.
---

Starting from the complete `ToothGrowth` data set, filter only the guinea pigs
that received ascorbic acid and had a dosage of 1.0mg or greater.

```{r tidy = FALSE, message = FALSE, warning = FALSE, error = FALSE}

# YOUR CODE HERE
tooth_growth_tibble = as_tibble(ToothGrowth)

tooth_growth_tibble %>% filter(supp == "VC", dose >= 1.0)


```



d.
---

Find out the mean length for the two supplement types. Do this with by
piping `tooth_growth_tibble` to `group_by()` and then to `summarise()`.  For
instructions read the help page for `summarise()`.

```{r tidy = FALSE, message = FALSE, warning = FALSE, error = FALSE}

# YOUR CODE HERE
tooth_growth_tibble %>% group_by(tooth_growth_tibble$supp) %>% summarise(meanlen = mean(tooth_growth_tibble$len))

```


**What is the mean tooth length for guinea pigs that received Orange Juice?**

*ANSWER:*
18.8133



___

# 3. Plotting data


a.
---

Using the `ToothGrowth` data set, create a plot of dose (x
axis) against length (y axis) using `ggplot()`. Show the two
supplement types in different colours.

```{r tidy = FALSE, message = FALSE, warning = FALSE, error = FALSE}

# YOUR CODE HERE
ggplot(data = ToothGrowth) + geom_point(aes(x= ToothGrowth$dose, y = ToothGrowth$len, colour = as.factor(supp)), size = 2, alpha = 0.8) + scale_x_continuous(name = "Dose", breaks = seq(0, 2, 0.5)) + scale_y_continuous(name = "Length", breaks = seq(0, 40, 5)) + ggtitle(label = "Tooth lengths based on dosage") + theme_bw()

```

**After visually inspecting the graph, does there seem to be any difference between the types of supplement?**

*ANSWER:*

With high dosages, the effects of the two supplements are almost similar; but in low dosages Orange Juice is more effective. 

___

End of homework sheet
